# EDA

## Map representation of distribution of properties

Here we decided to represent the distribution of properties in Switzerland using a map. The map is interactive and allows users to hover over the markers to see the price. The markers are color-coded in orange and have a semi-transparent fill to reduce visual noise. The size of the markers is smaller to optimize the visual representation of the data.

This visualization helps in understanding the geographical spread and density of properties in the dataset.

```{r}
# Create a leaflet map with optimized markers
map <- leaflet(properties_filtered) %>%
  addTiles() %>%  # Add default OpenStreetMap tiles
  addProviderTiles(providers$Esri.NatGeoWorldMap) %>% # Add topographic maps for context
  addCircleMarkers(
    ~lon, ~lat,
    radius = 1.5,  # Smaller radius for the circle markers
    color = "#F97300",  # Specifying a color for the markers
    fillOpacity = 0.2,  # Semi-transparent fill
    stroke = FALSE,  # No border to the circle markers to reduce visual noise
    popup = ~paste("Price: ", price, "<br>",
                   "Rooms: ", number_of_rooms, "<br>",
                   "Type: ", property_type, "<br>",
                   "Year: ", year_category),
    label = ~paste("Price: ", price)  # Tooltip on hover
  ) %>% addLegend(
    position = "bottomright",  # Position the legend at the bottom right
    colors = "#F97300",  # Use the same color as the markers
    labels = "Properties"  # Label for the legend
  )

map$width <- "100%"  # Set the width of the map to 100%
map$height <- 600  # Set the height of the map to 600 pixels

map
```

The map highlights that properties are well-distributed throughout the region, with fewer properties in the Alpine region, which is expected due to its mountainous terrain. We thus have a good representation of the data across different cantons and locations and we can use it for further analysis.

## Histogram of prices

```{r, warning=FALSE}
histogram_price <- ggplot(properties_filtered, aes(x = price)) +
  geom_histogram(binwidth = 100000, fill = "skyblue", color = "red") +
  labs(title = "Distribution of Prices",
       x = "Price",
       y = "Frequency") +
  theme_minimal()
# Convert ggplot object to plotly object
interactive_histogram_price <- ggplotly(histogram_price, width = 600, height = 400 )
# Display the interactive histogram
interactive_histogram_price
```

## Histogram of prices for each property type

note : only price between 0 and 500000 so some outliers aren't here

```{r, warning=FALSE}
# Create the ggplot object
histogram <- ggplot(properties_filtered, aes(x = price)) +
  geom_histogram(binwidth = 100000, fill = "skyblue", color = "black") +
  facet_wrap(~ property_type, scales = "free", ncol = 2) +
  labs(title = "Distribution of Prices by Property Type",
       x = "Price",
       y = "Frequency") +
  theme_minimal() +
  xlim(0, 5000000)

# Convert ggplot object to plotly object
interactive_histogram <- ggplotly(histogram, width = 600, height = 1000)

# Display the interactive plot
interactive_histogram

```

## Histogram of prices for each year category

note : only price between 0 and 500000 so some outliers aren't here

```{r, warning=FALSE}
# Create a histogram of prices for each year category
histogram <- ggplot(properties_filtered, aes(x = price)) +
  geom_histogram(binwidth = 100000, fill = "skyblue", color = "black") +
  facet_wrap(~ year_category, scales = "free", ncol = 2) +
  labs(title = "Distribution of Prices by Year Category",
       x = "Price",
       y = "Frequency") +
  theme_minimal() +
  xlim(0, 5000000)
# Convert ggplot object to plotly object
interactive_histogram_year <- ggplotly(histogram, width = 600, height = 1000)
# Display the interactive plot
interactive_histogram_year

```

## Histogram of prices for each canton

note : only price between 0 and 500000 so some outliers aren't here

```{r, warning=FALSE}
histogram <- ggplot(properties_filtered, aes(x = price)) +
  geom_histogram(binwidth = 100000, fill = "skyblue", color = "black") +
  facet_wrap(~ canton, scales = "free", ncol = 2) +
  labs(title = "Distribution of Prices by Canton for properties between 0 and 5 million",
       x = "Price",
       y = "Frequency") +
  theme_minimal() +
  xlim(0, 5000000)

# Convert ggplot object to plotly object with adjusted height
interactive_histogram <- ggplotly(histogram, width = 600, height = 1000) %>%
  layout(height = 1000)  # Adjust the height as needed

# Display the interactive plot
interactive_histogram

```
```{r, warning=FALSE}
histogram <- ggplot(properties_filtered, aes(x = price)) +
  geom_histogram(binwidth = 100000, fill = "skyblue", color = "black") +
  facet_wrap(~ canton, scales = "free", ncol = 2) +
  labs(title = "Distribution of Prices by Canton for properties between 5 million and 10 million",
       x = "Price",
       y = "Frequency") +
  theme_minimal() +
  xlim( 5000000,10000000)

# Convert ggplot object to plotly object with adjusted height
interactive_histogram <- ggplotly(histogram, width = 600, height = 1000) %>%
  layout(height = 1000)  # Adjust the height as needed

# Display the interactive plot
interactive_histogram

```
## Histogram of prices for each number of rooms

note : only price between 0 and 500000 so some outliers aren't here

```{r, warning=FALSE}
# Create a histogram of prices for each number of rooms
histogram <- ggplot(properties_filtered, aes(x = price)) +
  geom_histogram(binwidth = 100000, fill = "skyblue", color = "black") +
  facet_wrap(~ number_of_rooms, scales = "free", ncol = 2) +
  labs(title = "Distribution of Prices by Number of Rooms",
       x = "Price",
       y = "Frequency") +
  theme_minimal() +
  xlim(0, 5000000)

# Convert ggplot object to plotly object with adjusted height
interactive_histogram <- ggplotly(histogram, width = 600, height = 1000)%>%
  layout(height = 2000)

# Display the interactive plot
interactive_histogram

```
## Histogram of properties by square meters

```{r}
histogram <- ggplot(properties_filtered, aes(x = square_meters)) +
  geom_histogram(binwidth = 15, fill = "skyblue", color = "black") +
  labs(title = "Distribution of Properties by Square Meters",
       x = "Square Meters",
       y = "Frequency") +
  theme_minimal() +
  xlim(0, 2000)

# Convert ggplot object to plotly object with adjusted height
interactive_histogram <- ggplotly(histogram, width = NULL, height = NULL)  # Adjust width and height as needed

# Display the interactive plot
interactive_histogram

```

## Histogram of prices with impot

```{r, warning=FALSE}
# Create the boxplot
boxplot <- ggplot(properties_filtered, aes(x = as.factor(Tax_cluster), y = price)) +
  geom_boxplot(fill = "skyblue", color = "black") +
  labs(title = "Boxplot of Property Prices by Tax Cluster",
       x = "Tax Cluster",
       y = "Price") +
  theme_minimal() +
  ylim(100000, 400000)

# Convert ggplot object to plotly object
interactive_boxplot <- ggplotly(boxplot)
interactive_boxplot
```


```{r}
impot_cols <- names(properties_filtered)[startsWith(names(properties_filtered), "Impôt")]

# Count the number of NA values in selected columns
na_counts <- colSums(is.na(properties_filtered[impot_cols]))

# Print the counts
print(na_counts)
```

