# Unsupervised learning

* Clustering and/or dimension reduction

Trying to Cluster commune datas to:
1. Reduce dimension
2. See similarities

A regarder, est-ce qu'on fait un cluster pour les datas politques + un cluster pour les data démographiques, ou est-ce qu'on regroupe tout?

```{r}
set.seed(123)

# Clustering demographic
cols_commune_demographic <- select(commune, -c("REGION", "CODE_REGION","Conseil national - PLR","Conseil national - PDC", "Conseil national - PS", "Conseil national - UDC", "Conseil national - PEV/PCS", "Conseil national - PVL", "Conseil national - PBD", "Conseil national - PST/Sol.", "Conseil national - PES", "Conseil national - Petits partis de droite"))

# Scale the columns, some are total numbers, some are percentages
cols_commune_demographic <- scale(cols_commune_demographic)

# Calculate the distance matrix
dist_matrix_demographic <- dist(cols_commune_demographic, method = "minkowski")

# Perform hierarchical clustering
hclust_model_demographic <- hclust(dist_matrix_demographic, method = "ward.D")

# Create dendrogram
dend_demo <- as.dendrogram(hclust_model_demographic)
dend_demo <- color_branches(dend_demo, k = 5) #Set number of cluster to 5, to keep the same scale for all our variables

plot(dend_demo, main = "Demographics - Hierarchical Clustering Dendrogram")


```

```{r}
# Clustering politics
set.seed(123)

cols_commune_politics <- select(commune, c("Conseil national - PLR","Conseil national - PDC", "Conseil national - PS", "Conseil national - UDC", "Conseil national - PEV/PCS", "Conseil national - PVL", "Conseil national - PBD", "Conseil national - PST/Sol.", "Conseil national - PES", "Conseil national - Petits partis de droite"))

# Scale the columns, some are total numbers, some are percentages
cols_commune_politics <- scale(cols_commune_politics)

# Calculate the distance matrix
dist_matrix_politics <- dist(cols_commune_politics, method = "minkowski")

# Perform hierarchical clustering
hclust_model_politics <- hclust(dist_matrix_politics, method = "ward.D")

# Create dendrogram
dend_pol <- as.dendrogram(hclust_model_politics)
dend_pol <- color_branches(dend_pol, k = 5) #Set number of cluster to 5, to keep the same scale for all our variables

plot(dend_pol, main = "Politics - Hierarchical Clustering Dendrogram")


```

To prevent introducing 10 new types of taxes, we conducted a clustering analysis on the tax dataset to identify which municipalities can be grouped together. 
Based on the within-cluster sum of squares, we found 5 clusters. These 5 distinct clusters will be assigned to properties to determine which municipalities are subject to a particular type of tax.
## Tax
```{r}
set.seed(123)

# Clean data and convert to numeric
cleaned_impots <- apply(impots, 2, function(x) as.numeric(gsub("[^0-9.-]", "", x)))
cleaned_impots[is.na(cleaned_impots)] <- 0  # Replace NA values with 0

# Scale the features
scaled_impots <- scale(cleaned_impots)

# Perform k-means clustering
k <- 2  # Initial guess for the number of clusters
kmeans_model <- kmeans(scaled_impots, centers = k)

# Check within-cluster sum of squares (elbow method)
wss <- numeric(10)
for (i in 1:10) {
  kmeans_model <- kmeans(scaled_impots, centers = i)
  wss[i] <- sum(kmeans_model$withinss)
}
#plot(1:10, wss, type = "b", xlab = "Number of Clusters", ylab = "Within groups sum of squares")

# Adjust k based on elbow method
k <- 5  

# Perform k-means clustering again with optimal k
kmeans_model <- kmeans(scaled_impots, centers = k)

# Assign cluster labels to dendrogram
clusters <- kmeans_model$cluster

# Plot dendrogram
#colored_dend <- color_branches(dend, k = 5)
#y_zoom_range <- c(0, 80)  # Adjust the y-axis range as needed

#plot(colored_dend, main = "Hierarchical Clustering Dendrogram", horiz = FALSE, ylim = y_zoom_range)

```

```{r}
# Get the cluster centers
cluster_centers <- kmeans_model$centers

# Create a data frame with cluster centers
cluster_centers_df <- data.frame(cluster = 1:k, cluster_centers)

# Print cluster centers
print(cluster_centers_df)

# Calculate the size of each cluster
cluster_sizes <- table(kmeans_model$cluster)

# Print cluster sizes
print(cluster_sizes)

# Get the cluster labels
cluster_labels <- kmeans_model$cluster

# Convert cleaned_impots to a data frame
impots_cluster <- as.data.frame(cleaned_impots)

# Add the cluster labels to cleaned_impots
impots_cluster$cluster <- cluster_labels

rownames(impots_cluster) <- rownames(impots)

impots_cluster <- impots_cluster %>%
  rownames_to_column(var = "Community")

```

```{r}
# Preparing df_commune for merging with main dataset

df_commune <- select(commune, REGION)

df_commune$Demographic_cluster <- cutree(hclust_model_demographic, k = 5)
df_commune$Political_cluster <- cutree(hclust_model_politics, k = 5)

# Preparing to merge

merging <- inner_join(amto_df, df_commune, by = c("Community" = "REGION"))

impots_cluster_subset <- impots_cluster[, c("Community", "cluster")]
merging <- merging %>%
  left_join(impots_cluster_subset, by = "Community")

clusters_df <- merging %>%
  rename(Tax_cluster = cluster) %>%
  rename(Commune = Community)

clusters_df <- clusters_df %>%
  select(c("Commune", "zip_code", "Canton_code", "Demographic_cluster", "Political_cluster", "Tax_cluster"))

# Only NAs are for commune Brugg, (written Brugg (AG) in the other data set) -> j'entre le cluster à la mano
clusters_df$Tax_cluster[is.na(clusters_df$Tax_cluster)] <- 2

# adding it to our main data set:
properties_filtered <- merge(properties_filtered, clusters_df[, c("zip_code", "Demographic_cluster", "Political_cluster", "Tax_cluster")], by = "zip_code", all.x = TRUE)

na_count <- sum(is.na(properties_filtered[, c("Demographic_cluster", "Political_cluster", "Tax_cluster")]))

# Print the result
if (na_count > 0) {
  print("There are NA values in the merged dataframe.")
  print(na_count)
} else {
  print("There are no NA values in the merged dataframe.")
}

# Find rows with NA values in the specified columns
na_rows <- subset(properties_filtered, is.na(Demographic_cluster) | is.na(Political_cluster) | is.na(Tax_cluster))


```

